### 2023-06-06 guides/practical_guides.Rmd
practical guides

---
title: "Practical guides for implementing 'Five Pillar' principles"
author: "Mark Ziemann, Samuel S. Minot, Pierre Poulain, Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
bibliography: practical_guides.bib
csl: plos-computational-biology.csl
---

Mark Ziemann<sup>1*</sup>, Samuel S. Minot<sup>2</sup>, Pierre Poulain<sup>3</sup>,
Anusuiya Bora<sup>1</sup>

<details><summary>Author information (*click to expand*)</summary><p>

1. Deakin University, School of Life and Environmental Sciences, Geelong, Australia.

2. Fred Hutchinson Cancer Center, Seattle, USA.

3. Université Paris Cité, CNRS, Institut Jacques Monod, Paris, France.

Correspondance: m.ziemann@deakin.edu.au

| Author | ORCID |
| --- | --- |
| Mark Ziemann | 0000-0002-7688-6974 |
| Samuel S. Minot | 0000-0003-1639-3905 |
| Pierre Poulain | 0000-0003-4177-3619 |
| Anusuiya Bora | 0009-0006-2908-1352 |

</details>
</p>
Source: https://github.com/markziemann/5pillars/blob/main/manuscript/5pillars.Rmd

## Practical guides

The five pillar approach is a synthesis of over a decade of learnings around computational
reproducibility, and contains six key recommendaions and many best practices.
These may be overwhelming for any data scientists, especially beginners and novices,
so here we have tried to put together some recommended resources to enable practitioners to put these
principles into practice.
One of the questions practitioners may have is *"Where do I start?"* and
*"In what order should I learn/master these principles?"*.
The following sections have been arranged in order, and have been submitted to the
[Internet Archive](https://archive.org/) for posterity.

### Getting started with data analysis in R and Python

Learn the basics of your scripting language (R/Python/shell) so that you can do useful analyses
with it. Learn by solving problems.

* Intro to R and Rstudio

* Intro to Python and iPython IDE

* Intro to vscode

### Literate programming with R Markdown, Jupyter and Quarto

Convert those scripts into literate scripts and document them with introduction, methods,
code comments, results observations and interpretations.
Use JupyterLab (Jupyter notebook) or Rstudio (Rmarkdown) development environments to provide
graphical interfaces to the programming languages.

* https://medium.com/@itsjzt/beginner-guide-to-markdown-229adce30074

* https://www.dataquest.io/blog/r-markdown-guide-cheatsheet/

* https://tysonbarrett.com/Rstats/chapter-9-reproducible-workflow-with-rmarkdown.html

* https://andrewbtran.github.io/NICAR/2017/reproducibility-rmarkdown/rmarkdown.html

* https://quarto.org/docs/computations/r.html

* https://quarto.org/docs/computations/python.html

* https://jupyter-notebook-beginner-guide.readthedocs.io/en/latest/

* https://www.dataquest.io/blog/jupyter-notebook-tutorial/

* https://medium.com/velotio-perspectives/the-ultimate-beginners-guide-to-jupyter-notebooks-6b00846ed2af

### Practical guides for `git` and GitHub

Learn how integrate version control into your development environment.

* https://www.datacamp.com/tutorial/github-and-git-tutorial-for-beginners/

* https://www.freecodecamp.org/news/introduction-to-git-and-github/

* https://www.geo.uzh.ch/microsite/reproducible_research/post/rr-rstudio-git/

* https://anderfernandez.com/en/blog/how-to-link-and-use-git-with-rstudio/

* https://andgarc.github.io/basic-git-jupyter-lesson/index.html

* https://blog.reviewnb.com/github-jupyter-notebook/

* https://andgarc.github.io/basic-git-jupyter-lesson/index.html

### Practical guides to data sharing

Learn about the FAIR principles and make the research data available in a public repository.

### Extend the scripts to make them end-to-end processes

Try to extend the scripts to make them end-to-end processes, by linking to raw data and outputting
whole figures or research articles.

### Practical guides for Conda, Guix and Docker

Control the environment.
Record the version of programming languages and packages.
Become familiar and experiment with Conda, Guix and Docker.
Select the option that works best for your project.

### Practical guides for documenting computational research

Document the process of reproduction as if you were an external researcher.
Test the instructions.

### Moving from "dev" to "prod": testing and continuous analysis

Test the reproduction process as if you were an external researcher.